Variants of G\"{o}del's Ontological Proof \\ in a Natural Deduction Calculus

### 2016-09-08 Papers/InProgress/NaturalDeduction/GodProof-ND.tex
Reviewer 2, comment A (i).
Reviewer 2: Leibniz's Equality

\begin{definition} \emph{Terms and formulas} are given by the following grammar:
\begin{align*}
 s,t \quad ::= \quad & 
  p_\tau \ \mid \ 
  X_\tau \ \mid \
  (\lambda X_\theta.s_\tau)_{\theta\imp\tau} \ \mid \ 
  (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
  & s_{\tau} = t_{\tau} \ \mid \\
& \bot_o \ \mid \
  \imp_{o\imp o\imp o} \ \mid \ 
  \wedge_{o\imp o\imp o} \ \mid \\
& \all_{(\tau\imp o)\imp o} \ \mid \ 
  \ex_{(\tau\imp o)\imp o} \ \mid \
  \nec_{o\imp o} \ \mid \
  \pos_{o\imp o}
\end{align*}
where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$. Parenthesis conventions, infix notation for propositional connectives and binding notation for quantifiers are assumed. Furthermore, subscript types are omitted when they are clear from the context. Inequality ($\neq_{\tau\imp\tau\imp o}$), negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$) and equivalence ($\biimp_{o\imp o\imp o}$) are defined by $s \neq t\equiv \neg (s=t)$, $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$ and $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$.
\end{definition}

The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}. The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}. Finally, modal operators are handled by special rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
%Apart from the use of labels and the dual rules for `$\pos$', these rules are essentially the modal rules from \cite{todo}. 
Beta-reduction is implicit; all rules are assumed to operate modulo beta-reduction. A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules. A \emph{proof} of a formula $F$ is a derivation without open assumptions and having $F$ as root not inside any box.

@@

The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical}, \ref{fig:QuantifierRules} and \ref{fig:EqualityRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.

>>>>

\begin{definition} \emph{Terms and formulas} are given by the following grammar:
\begin{align*}
 s,t \quad ::= \quad & 
  p_\tau \ \mid \ 
  X_\tau \ \mid \
  (\lambda X_\theta.s_\tau)_{\theta\imp\tau} \ \mid \ 
  (s_{\theta\imp\tau}\, t_\theta)_\tau \ \mid \\
& \bot_o \ \mid \
  \imp_{o\imp o\imp o} \ \mid \ 
  \wedge_{o\imp o\imp o} \ \mid \\
& \all_{(\tau\imp o)\imp o} \ \mid \ 
  \ex_{(\tau\imp o)\imp o} \ \mid \
  \nec_{o\imp o} \ \mid \
  \pos_{o\imp o}
\end{align*}
where $p_\tau$ and $X_\tau$ range over, respectively, constants and variables of type $\tau$. Parenthesis conventions, infix notation for propositional connectives and binding notation for quantifiers are assumed. Furthermore, subscript types are omitted when they are clear from the context. Negation ($\neg_{o\imp o}$), disjunction ($\vee_{o\imp o\imp o}$), equivalence ($\biimp_{o\imp o\imp o}$), equality ($=_{\tau\imp\tau\imp o}$), inequality ($\neq_{\tau\imp\tau\imp o}$), are defined by $\neg A\equiv A\imp \bot$, $(A \vee B) \equiv \neg (\neg A \wedge \neg B)$, $ (A\biimp B)\equiv (A\imp B)\wedge (B\imp A)$, $s = t \equiv \forall P. (P \ s) \biimp (P \ t)$ (Leibniz's equality) and $s \neq t\equiv \neg (s=t)$, .
\end{definition}

The natural deduction calculus used here has standard rules for propositional connectives and quantifiers, as shown in Figures \ref{fig:PropositionalRules} and \ref{fig:QuantifierRules}. The extension to classical logic is achieved by adding a rule for double negation elimination, shown in Figure \ref{fig:Classical}. Finally, modal operators are handled by special rules that insert or remove formulas from boxes, as shown in Figure \ref{fig:NDK}.
%Apart from the use of labels and the dual rules for `$\pos$', these rules are essentially the modal rules from \cite{todo}. 
Beta-reduction and alpha-conversion are implicit: all rules are assumed to operate modulo these operations. A \emph{derivation} is a directed acyclic graph whose nodes are formulas and whose edges correspond to applications of the inference rules. A \emph{proof} of a formula $F$ is a derivation without open assumptions and having $F$ as root not inside any box.

@@

The calculus having only the rules shown in Figures \ref{fig:PropositionalRules}, \ref{fig:Classical} and \ref{fig:QuantifierRules} is named \ND, while the calculus with the additional rules shown in Figure \ref{fig:NDK} is named \NDK.
